Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2022-09-08 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

```{r setup, include=FALSE}
library(readxl)
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
source("~/GitHub/CF/get_genotype_class_severity.R")
```

```{r}
# Import exclusion data from Excel
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude)]
# Import data
df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
# Remove exclusions from Angela
df <- df[!(df$PatientID %in% exclude), ]
df$Diagnosis <- diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)]
# Filter per Maxie's code
df <- df %>%
  filter(
    Age.at.Test.in.Years < 50,
    !(Diagnosis == "Non-CF" & PatientID %in%
      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793))
  ) %>%
  distinct(.) %>%
  filter(Sex != "", Demographics.Available == "TRUE") %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7)
# Add modulator info
cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator <- apply(df, 1, function(r) {
  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
  if (any(duplicated(mod$Age.at.Start..years.))) {
    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
  }
  mod <- mod[order(mod$Age.at.Start..years.), ]
  age <- as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
    labels = c("None", mod$Modulator), right = F
  ))
})
# Genotype info
df$geno1 <- get_cf_genotype_class_severity(df$Genotypes1)
df$geno2 <- get_cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno1, df$geno2, sep = "_")
df$geno_risk <- factor(df$geno_risk,
  levels = c(
    "Mild_Severe", "Mild_Unknown", "Not CF_Not CF",
    "Severe_Mild", "Mild_Mild",
    "Severe_Severe", "Severe_Unknown", "Unknown_Severe",
    "Unknown_Unknown"
  ),
  labels = c(
    "Low", "Low", "Not CF", "Low", "Low", "High", "Unknown",
    "Unknown", "Unknown"
  )
)
# Merge left and right values
df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
  amt <- suppressWarnings(as.numeric(as.character(r)))
  mean(amt, na.rm = T)
})
df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
  cl <- suppressWarnings(as.numeric(as.character(r)))
  mean(cl, na.rm = T)
})
df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
  na <- suppressWarnings(as.numeric(as.character(r)))
  mean(na, na.rm = T)
})
df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
  k <- suppressWarnings(as.numeric(as.character(r)))
  mean(k, na.rm = T)
})
# Sort
df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
```

# Table 1: Patient Demographics and Sample Characteristics

@@

```{r results='asis'}
t1_df <- df %>%
  filter(Age.at.Test.in.Days >= 2) %>%
  group_by(PatientID) %>%
  arrange(Age.at.Test.in.Weeks) %>%
  filter(row_number() == 1)
t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + geno_risk + Amount + Cl + Na + K,
  data = t1_df
)
summary(t1,
  labelTranslations =
    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
)
```

@@

## Table

@@

## Plot

```{r}
ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw()
```

```{r}
# Write only those with multiple values
mult <- unique(df$PatientID[which(duplicated(df$PatientID))])
multiples <- df[df$PatientID %in% mult, ]
write.csv(multiples, file = "./Data_Cleaned/multiple_values.csv", na = "", row.names = F)
```

>>>>

```{r setup, include=FALSE}
library(readxl)
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
source("~/GitHub/CF/get_genotype_class_severity.R")
```

```{r}
# Import exclusion data from Excel
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude)]
# Import data
df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
# Remove exclusions from Angela
df <- df[!(df$PatientID %in% exclude), ]
df$Diagnosis <- diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)]
# Filter per Maxie's code
df <- df %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7) %>%
  filter(
    Age.at.Test.in.Years < 50 & Age.at.Test.in.Days >= 2,
    !(Diagnosis == "Non-CF" & PatientID %in%
      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793))
  ) %>%
  distinct(.) %>%
  filter(Sex != "", Demographics.Available == "TRUE")
# Add modulator info
cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator <- apply(df, 1, function(r) {
  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
  if (any(duplicated(mod$Age.at.Start..years.))) {
    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
  }
  mod <- mod[order(mod$Age.at.Start..years.), ]
  age <- as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
    labels = c("None", mod$Modulator), right = F
  ))
})
# Genotype info
df$geno1 <- get_cf_genotype_class_severity(df$Genotypes1)
df$geno2 <- get_cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno1, df$geno2, sep = "_")
df$geno_risk <- factor(df$geno_risk,
  levels = c(
    "Mild_Severe", "Mild_Unknown", "Not CF_Not CF",
    "Severe_Mild", "Mild_Mild",
    "Severe_Severe", "Severe_Unknown", "Unknown_Severe",
    "Unknown_Unknown"
  ),
  labels = c(
    "Low", "Low", "Not CF", "Low", "Low", "High", "Unknown",
    "Unknown", "Unknown"
  )
)
# Which arm
df$Arm <- apply(df[, c("Cl", "Cl_Left", "Cl_Right", "Na", "Na_Left", "Na_Right", "K", "K_Left", "K_Right")], 1, function(r) {
  k <- sum(!is.na(r))
  if (k == 6) {
    return("Both")
  } else if (k == 3) {
    return("Arm Not Noted")
  } else {
    return(NA)
  }
})
# Merge left and right values
df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
  amt <- suppressWarnings(as.numeric(as.character(r)))
  mean(amt, na.rm = T)
})
df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
  cl <- suppressWarnings(as.numeric(as.character(r)))
  mean(cl, na.rm = T)
})
df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
  na <- suppressWarnings(as.numeric(as.character(r)))
  mean(na, na.rm = T)
})
df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
  k <- suppressWarnings(as.numeric(as.character(r)))
  mean(k, na.rm = T)
})
# Sort
df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
# Get list of those multiple values
mult <- unique(df$PatientID[which(duplicated(df$PatientID))])
```

# Patient Demographics and Sample Characteristics

@@

## By group

```{r results='asis'}
t1_df <- df %>%
  group_by(PatientID) %>%
  arrange(Age.at.Test.in.Weeks) %>%
  filter(row_number() == 1) # First measure for everyone
t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + geno_risk + Amount + Cl + Na + K,
  data = t1_df
)
summary(t1,
  labelTranslations =
    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
)
```

## By Method

### All participants

```{r results='asis'}
t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + Arm,
  data = t1_df
)
summary(t1,
  labelTranslations =
    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
)
```

### Participants with more than one measure

```{r results='asis'}
t1 <- tableby(Method ~ Sex + Age.at.Test.in.Years + Diagnosis + geno_risk + Amount + Cl + Na + K + Arm,
  data = t1_df[t1_df$PatientID %in% mult, ]
)
summary(t1,
  labelTranslations =
    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
)
```

@@

@@

# Plots

```{r warning=FALSE}
ggplot(df, aes(x = Diagnosis, y = Na)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Sweat Sodium")

ggplot(df, aes(x = Diagnosis, y = K)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Sweat Potassium")

ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Sweat Chloride")
```

```{r}
# Write only those with multiple values
multiples <- df[df$PatientID %in% mult, ]
write.csv(multiples, file = "./Data_Cleaned/multiple_values.csv", na = "", row.names = F)
```
